---
title: "Proof of Concept Interaction Analysis"
author: "Nathan Strozewski"
date: "2023-05-17"
output: html_document
---

```{r libraries, message=FALSE}

# run these first

library(ggplot2) # plotting
library(ggrepel) # plot point labels
library(dplyr) # data wrangling
library(readxl) # read file from OneDrive
library(tidyverse)

```

```{r set custom theme}

custom_theme <- function() { 
  font <- "Helvetica" # font selection
    
    theme_minimal() %+replace% # theme based on minimal with following replacements
    
    theme(panel.grid.major = element_blank(), # leave grids and axis ticks blank
          panel.grid.minor = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_line(color = "black",
                                   size = 1),
          panel.border = element_rect(color = "black",
                                      fill=NA,
                                      size=1),
          plot.title = element_text(family = font,
                                    size = 20,
                                    face = 'bold',
                                    hjust = 0.5, # move title to center horizontally
                                    vjust = 2), # move title up a wee bit
          plot.subtitle = element_text(family = font,
                                       size = 15,
                                       hjust = 0.5),
          plot.caption = element_text(family = font,
                                      size = 10,
                                      hjust = 1), # put caption in right corner
          axis.title = element_text(family = font,
                                    face = 'italic',
                                    size = 15),
          axis.text = element_text(family = font,
                                   size = 10),
          axis.text.x = element_text(margin = margin(t = 2, # top
                                                     r = 2, # right
                                                     b = 2, # bottom
                                                     l = 2))) # left
}

```

The file is not accessible on OneDrive, so I had to download and store locally on my computer in the project's working directory. Future replication needs to update this file pathway:

```{r get data}

all_apms <- read_xlsx("A002_B001_C005_D003_SAINT Report.xlsx") %>% # read data
  rename('bait' = ait) %>% # fix typo in data file
  select(-SpecSum, -AvgSpec, -NumReplicates, -boosted_by) %>% # remove unnecessary cols
  mutate(bait = recode(bait, DrosophilaCicGFP = 'Drosophila Cic-GFP', HumanCicGFP = 'Human Cic-GFP')) %>%  # rename samples for easier plotting
  mutate('log2FoldChange' = log2(FoldChange)) %>% # calculate log of fold change
  mutate('neglog10AvgP' = -log10(AvgP)) %>% # calculate log of probability
  mutate('neglog10BFDR' = -log10(BFDR))

all_human <- all_apms %>% 
  filter(bait == 'Human') %>% 
  select(-bait) # human specific data set

all_fly <- all_apms %>% 
  filter(bait == 'Drosophila') %>% 
  select(-bait) # fly specific data set

filtered_apms <- all_apms %>% 
  arrange(bait, PreyGene) %>% 
  filter(AvgP > 0.79,
         ctrlCounts == 0) # 

```

Now that the data is ready, I'll visualize to see what's going on. I'll start by looking at all the hits that are also present in the control. Fold Change will be the indicator of interaction strength in experimental compared to control, while Bayesian false discovery rate (BFDR) will be the indicator of statistical significance or likelihood that the interaction is not a false discovery. I'll convert these to log form, which is standard across the literature for these analyses. This analysis below will indicate which interactions, that are also present in the control, are stronger hits for human and fly Cic.

```{r initial volcano visualization}

ubiquitous_apms <- all_apms %>% 
  # filter(ctrlCounts > 0) %>% 
  mutate(Color = if_else(neglog10BFDR > 1.3 & log2FoldChange > 1.3,
                         "red",
                         if_else(neglog10BFDR < 1.3 & log2FoldChange > 1.3,
                                 "blue",
                                 "grey"))) %>% # for colors on geom_point() to correspond to certain graph placements
  mutate(repel_label = ifelse(Color %in% c("red", "blue"),
                              PreyGene,
                              "")) # for labels in geom_text_repel() later

ubiquitous_volcano <- ggplot(data = ubiquitous_apms,
                             mapping = aes(x = log2FoldChange,
                                           y = neglog10BFDR,
                                           color = Color)) +
  geom_point() +
  geom_jitter() + # separate the points a bit
  scale_color_manual(values = c("blue", "grey", "red")) + # sets colors
  custom_theme() + # my custom theme to make things look nice
  theme(legend.position = "none") + # remove legend
  facet_wrap(vars(bait)) + # split by human and drosophila
  geom_hline(yintercept = 1.3,
             linetype = 3,
             alpha = 0.5) + # horizontal line for acceptable Fold Change
  geom_vline(xintercept = 1.3,
             linetype = 3,
             alpha = 0.5) + # vertical line for acceptable probability
  # geom_text_repel(aes(label = repel_label), # sets labels to the repel_label col
  #                 force = 2, # how far label should be from point
  #                 size = 2.5, # size of text
  #                 box.padding = 0.33, # space around label
  #                 max.overlaps = 100) + # how many overlaps can be processed
  coord_cartesian(xlim = c(-4, 4),
                  ylim = c(0, 2)) +
  scale_x_continuous(breaks = c(-3, -2, -1, 0, 1, 2, 3)) + # change x axis
  scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2)) + # change y axis
  labs(caption = "n = 1",
       x = "log2(Fold Change)",
       y = "-log10(FDR)")

ubiquitous_volcano

ggsave("Logarithmic Analysis of AP-MS Data.pdf",
       height = 5,
       width = 8)
ggsave("Volcano, Interactors Present in Control.pdf",
       height = 10,
       width = 14)

```

Now, let's look at the hits that are unique only to human and fly Cic, and are not hits in the control. These results will potentially be clearer indicators of strong interaction. The only key component of the code I have to change is in the dataset - setting the filter to ctrlCounts == 0 instead of > 0 will make it so only hits that are unique to the experimental conditions will be included. This graph will look much different - since there are no hits in the control for any of these hits, 

```{r initial volcano visualization}

experimental_only_apms <- all_apms %>% 
  filter(ctrlCounts == 0) %>% # only hits unique to experimental condition will be included
  mutate(Color = if_else(neglog10BFDR > 1.3 & log2FoldChange > 1.3,
                         "red",
                         if_else(neglog10BFDR < 1.3 & log2FoldChange > 1.3,
                                 "blue",
                                 "grey"))) %>% # for colors on geom_point() to correspond to certain graph placements
  mutate(repel_label = ifelse(Color %in% c("red", "blue"),
                              PreyGene,
                              "")) # for labels in geom_text_repel() later

ubiquitous_volcano <- ggplot(data = experimental_only_apms,
                             mapping = aes(x = log2FoldChange,
                                           y = neglog10BFDR,
                                           color = Color)) +
  geom_point() +
  geom_jitter() + # separate the points a bit
  scale_color_manual(values = c("blue", "red", "grey")) + # sets colors
  custom_theme() + # my custom theme to make things look nice
  theme(legend.position = "none") + # remove legend
  facet_wrap(vars(bait)) + # split by human and drosophila
  geom_hline(yintercept = 1.3,
             linetype = 3,
             alpha = 0.5) + # horizontal line for acceptable Fold Change
  geom_vline(xintercept = 1.3,
             linetype = 3,
             alpha = 0.5) + # vertical line for acceptable probability
  # geom_text_repel(aes(label = repel_label), # sets labels to the repel_label col
  #                 force = 3, # how far label should be from point
  #                 size = 2, # size of text
  #                 box.padding = 0.5, # space around label
  #                 max.overlaps = 1000) + # how many overlaps can be processed
  coord_cartesian(xlim = c(0, 10),
                  ylim = c(0, 2)) +
  scale_x_continuous(breaks = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10)) + # change x axis
  scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2)) + # change y axis
  labs(caption = "n = 1",
       x = "log2(Fold Change)",
       y = "-log10(p)")

ubiquitous_volcano
ggsave("Volcano, Interactors Unique to Experimental Conditions.pdf",
       height = 10,
       width = 14)

```

Alright, let's put a table together to highlight some of this. I'll make a table for all the strong, significant hits for both organisms - split into two tables by whether or not the hits were also present in the control.

```{r subsetting the data}

ubiquitous_fly <- all_apms %>% 
  filter(bait == "Drosophila",
         ctrlCounts != 0,
         neglog10BFDR > 1.3,
         log2FoldChange > 1.3)

experimental_fly <- all_apms %>% 
  filter(bait == "Drosophila",
         ctrlCounts == 0,
         neglog10BFDR > 1.3,
         log2FoldChange > 1.3)

ubiquitous_human <- all_apms %>% 
  filter(bait == "Human",
         ctrlCounts != 0,
         neglog10BFDR > 1.3,
         log2FoldChange > 1.3)

experimental_human <- all_apms %>% 
  filter(bait == "Human",
         ctrlCounts == 0,
         neglog10BFDR > 1.3,
         log2FoldChange > 1.3)

ubiquitous_all_table <- all_apms %>% 
  filter(ctrlCounts != 0,
         neglog10BFDR > 1.3,
         log2FoldChange > 1.3) %>% 
  arrange(PreyGene, bait, "-log10p") %>% 
  mutate("-log10(p)" = neglog10BFDR,
         "log2(Fold Change)" = log2FoldChange) %>% 
  select(bait, PreyGene, Spec, ctrlCounts, "-log10(p)", "log2(Fold Change)", Prey) %>% 
  rename("Bait" = bait,
         "Gene" = PreyGene,
         "Hits" = Spec,
         "Control Hits" = ctrlCounts,
         "UniProt ID" = Prey) %>% 
  as.data.frame() %>% 
  write_csv("Strong, Significant Interactors that are Present in the Control.csv")

experimental_all_table <- all_apms %>% 
  filter(ctrlCounts == 0,
         neglog10BFDR > 1.3,
         log2FoldChange > 1.3) %>% 
  arrange(PreyGene, bait, "-log10p") %>% 
  mutate("-log10(p)" = neglog10BFDR,
         "log2(Fold Change)" = log2FoldChange) %>% 
  select(bait, PreyGene, Spec, ctrlCounts, "-log10(p)", "log2(Fold Change)", Prey) %>% 
  rename("Bait" = bait,
         "Gene" = PreyGene,
         "Hits" = Spec,
         "Control Hits" = ctrlCounts,
         "UniProt ID" = Prey) %>% 
  as.data.frame() %>% 
  write_csv("Strong, Significant Interactors that are Absent in the Control.csv")

```








