---
title: "Cic Phospho-mutant Neurobehavioral Assays"
author: "Nathan Strozewski"
date: "2024-03-29"
output: html_document
---


## Libraries

```{r libraries}

library(dplyr)
library(tibble)
library(data.table)
library(ggplot2)
library(tidyr)
library(readxl)
library(ggpubr)
library(devtools)
library(gridExtra)
library(performance)

```

## Custom Theme

```{r custom_theme and color palette}
custom_theme <- function() { 
  font <- "Helvetica" # font selection
    
    theme_minimal() %+replace% # theme based on minimal with following replacements
    
    theme(panel.grid.major = element_blank(), # leave grids and axis ticks blank
          panel.grid.minor = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_line(color = "black",
                                   size = 1),
          panel.border = element_rect(color = "black",
                                      fill=NA,
                                      size=1),
          plot.title = element_text(family = font,
                                    size = 20,
                                    face = 'bold',
                                    hjust = 0.5, # move title to center horizontally
                                    vjust = 2), # move title up a wee bit
          plot.subtitle = element_text(family = font,
                                       size = 15,
                                       hjust = 0.5),
          plot.caption = element_text(family = font,
                                      size = 10,
                                      hjust = 1), # put caption in right corner
          axis.title = element_text(family = font,
                                    size = 15),
          axis.text = element_text(family = font,
                                   size = 10),
          axis.text.x = element_text(margin = margin(t = 2, # top
                                                     r = 2, # right
                                                     b = 2, # bottom
                                                     l = 2))) # left
}

custom_palette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

```

## Load aggregated data file

```{r load_data}

geo_data <- read_xlsx('Geotaxis_Elav x Cic Phosphomutants_2024_01_24.xlsx') %>% 
            mutate(distance_cm = distance_in * 2.54,
                   relative_dist = distance_cm / 15.24) %>% 
            mutate(genotype = recode(genotype, "elav>GFP" = 'GFP',
                            "elav>Cic WT-mV" = 'Cic WT-mV',
                            "elav>Cic 6A-mV" = 'Cic 6A-mV'))

photo_data <- read_xlsx('Phototaxis_Elav x Cic Phosphomutants_2024_01_24.xlsx') %>% 
            mutate(distance_cm = distance_in * 2.54,
                   relative_dist = distance_cm / 15.24) %>% 
            mutate(genotype = recode(genotype, "elav>GFP" = 'GFP',
                            "elav>Cic WT-mV" = 'Cic WT-mV',
                            "elav>Cic 6A-mV" = 'Cic 6A-mV'))


```

## Geo Data Summary 

```{r geo_summary}

geo_summary <- geo_data %>% 
               group_by(genotype, time_sec) %>%  
               mutate(genotype = factor(genotype, levels = c('GFP', 'Cic WT-mV', 'Cic 6A-mV'))) %>% 
               summarise(mean_rel_dist = mean(relative_dist),
                         sd_rel_dist = sd(relative_dist),
                         upper = mean_rel_dist + sd_rel_dist,
                         lower = mean_rel_dist - sd_rel_dist)

geo_summary

geo_summary_factors <- geo_summary %>% 
  mutate(time_sec = as.factor(time_sec))

geo_aov <- aov(mean_rel_dist ~ genotype + time_sec,
               data = geo_summary_factors)
summary(geo_aov)

geo_geno_tukeyHSD <- TukeyHSD(geo_aov, "genotype")
print(geo_geno_tukeyHSD)

```
No statistical impact of time on distance. All genotypes were statistically distinct.

```{r basic_geo_viz}

geo_plot <- ggplot(data = geo_summary,
                   mapping = aes(x = time_sec,
                                 y = mean_rel_dist,
                                 group = genotype)) +
  geom_line(aes(color = genotype,
                linetype = genotype),
            stat = "summary") +
  geom_errorbar(aes(ymin = lower,
                    ymax = upper),
                alpha = 0.75, size = 0.2, width = 0.25,
                position = "dodge") +
  scale_shape(guide = FALSE) +
  scale_linetype(guide = FALSE) +
  scale_color_manual(values = c("#000000", "#56B4E9", "#D55E00")) +
  custom_theme() +
  theme(legend.title = element_text(hjust = 0.5, face = "bold", size=14),
        legend.text=element_text(size=12),
        plot.subtitle = element_text(hjust = 0, size = 12)) +
  scale_x_continuous(breaks = c(0, 5, 10, 15, 20, 25, 30)) +
  scale_y_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Time (sec)",
       y = "Distance Traveled Upwards",
       color = "Genotype",
       linetype = "Genotype",
       subtitle = expression(paste("ANOVA: ", italic("F"), "= 306.9, ", italic("p "), "= <0.001")))

geo_plot

ggsave("Figure 7B_elav-GAL4 Geotaxis.pdf",
       width = 3000,
       height = 1500,
       units = "px")

```

```{r geo_summary}

photo_summary <- photo_data %>% 
               group_by(genotype, time_sec) %>%  
               mutate(genotype = factor(genotype, levels = c('GFP', 'Cic WT-mV', 'Cic 6A-mV'))) %>% 
               summarise(mean_rel_dist = mean(relative_dist),
                         sd_rel_dist = sd(relative_dist),
                         upper = mean_rel_dist + sd_rel_dist,
                         lower = mean_rel_dist - sd_rel_dist)

photo_summary

photo_summary_factors <- photo_summary %>% 
  mutate(time_sec = as.factor(time_sec))

photo_aov <- aov(mean_rel_dist ~ genotype + time_sec,
               data = photo_summary_factors)
summary(photo_aov)

photo_geno_tukeyHSD <- TukeyHSD(photo_aov, "genotype")
print(photo_geno_tukeyHSD)

```

```

```{r basic_geo_viz}

photo_plot <- ggplot(data = photo_summary,
                   mapping = aes(x = time_sec,
                                 y = mean_rel_dist,
                                 group = genotype)) +
  geom_line(aes(color = genotype,
                linetype = genotype),
            stat = "summary") +
  geom_errorbar(aes(ymin = lower,
                    ymax = upper),
                alpha = 0.75, size = 0.2, width = 0.25,
                position = "dodge") +
  scale_shape(guide = FALSE) +
  scale_linetype(guide = FALSE) +
  scale_color_manual(values = c("#000000", "#56B4E9", "#D55E00")) +
  custom_theme() +
  theme(legend.title = element_text(hjust = 0.5, face = "bold", size=14),
        legend.text=element_text(size=12),
        plot.subtitle = element_text(hjust = 0, size = 12)) +
  scale_x_continuous(breaks = c(0, 5, 10, 15, 20, 25, 30)) +
  scale_y_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  labs(x = "Time (sec)",
       y = "Distance Traveled Towards Light",
       color = "Genotype",
       linetype = "Genotype",
       subtitle = expression(paste("ANOVA: ", italic("F"), "= 833.2, ", italic("p "), "= <0.001")))

photo_plot

ggsave("Figure 7A_elav-GAL4 Phototaxis.pdf",
       width = 3000,
       height = 1500,
       units = "px")

```

